---
title: "Field Data Reburn points"
author: "Anna Talucci"
date: "2024-03-18"
output: html_document
---

# Overview
Organize field data to select for burn_depth and recent burns from 2016 to present. The goal is to pair these sites with SAR data.

Field data provided by Xanthe Walker from FiSL and Legacy Carbon Project. 

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data
```{r}
lcsites = read_csv("../data/fieldCsv/LC_SITES_11142023.csv")
lcdata = read_csv("../data/fieldCsv/LC_Data_09052023.csv")
fd = read_csv("../data/fieldCsv/FiSL_sitedata_BB2023.csv")
fp = read_csv("../data/fieldCsv/FINAL_PLOTS_01142024.csv")
```

```{r}
lcsites
lcdata
fp
```

above ground combustion trees + shrubs + cwd
below ground 

```{r}
fp %>% filter(unique_site_id.x==117 | unique_site_id.x ==174)
```
# Filter Data for years of interest

## For final plot data
Notes: 
- Projection/Datum NAD83
- by select only fire from 2016 on we only loos 96 sites
- burn_depth has 197 NA

```{r}
fp[is.na(fp$burn_depth),]
```
```{r}
( fp1 = fp %>% 
    dplyr::select(unique_site_id.x, date.x, official_fire_scar, site, latitude, longitude, moisture_class, moisture_numeric, recent_burn, second_burn, triple_burn, reburn_age, reburn_class, prefire.trees:burn_depth) %>%
    rename(siteID = unique_site_id.x, date=date.x) %>%
    filter(recent_burn >=2016) %>%
    drop_na(burn_depth)
  )
```

## Reburn sites
```{r}
reburn = fp %>% 
    dplyr::select(unique_site_id.x, date.x, official_fire_scar, site, latitude, longitude, moisture_class, moisture_numeric, recent_burn, second_burn, triple_burn, reburn_age, reburn_class, prefire.trees:burn_depth) %>%
    rename(siteID = unique_site_id.x, date=date.x) %>%
  drop_na(second_burn)
```

```{r}
unique(sort(reburn$recent_burn))
```
## For Legacy Carbon (LC) data
Notes: 
- Projection/Datum NAD83 
- by select only fire from 2016 on we only loos 96 sites
- burn_depth has 58 NA
```{r}
lcdata
```
```{r}
lcdata[is.na(lcdata$Burn_Depth),]
```
```{r}
( lc1 = lcdata %>% 
    dplyr::select(Site_Name, Sampling_Date, Year_of_Burn, Latitude, Longitude, Moisture_Class,  SOL_depth, Burn_Depth) %>%
    rename(burn_depth = Burn_Depth, date=Sampling_Date, recent_burn=Year_of_Burn, latitude=Latitude, longitude=Longitude,) %>%
    filter(recent_burn >=2016) %>%
    drop_na(burn_depth) %>%
  drop_na(latitude)
  )
```


# To point shp
```{r}
fp1_pts = st_as_sf(fp1, coords = c("longitude","latitude"), crs = 4269, remove = FALSE)
```

```{r}
lc1_pts = st_as_sf(lc1, coords = c("longitude","latitude"), crs = 4269, remove = FALSE)
```

```{r}
reburn_pts = st_as_sf(reburn, coords = c("longitude","latitude"), crs = 4269, remove = FALSE)
```
# Reproject to WSG 84 
4326
```{r}
reburn4326 = st_transform(reburn_pts, crs=4326)
```

### write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(fp1_pts, "../outputs/fieldData/FiSLBurnDepth.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(reburn4326, "../outputs/spatialDataCreate/fieldData/FiSLReburn4326.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(lc1_pts, "../outputs/fieldData/LcBurnDepth.shp", driver="ESRI Shapefile")
```