---
title: "LightningExploreData"
author: "Anna Talucci"
date: "2024-02-15"
output: html_document
---

# Overview

Explore Alaska Lightning data. Data for download within the zipped files on the [site](https://fire.ak.blm.gov/predsvcs/maps.php).

Lightning data to be used by Xanthe to evaluate wildfire probability. 

Notes:
-- Temporal period: 2013 to 2022
-- Spatial domain: Ecoregions of AK boreal interior, boreal cordillera, taiga plains, and taiga shield.
-- Lightning : yes I think we want it as precise as possible.  However, in the proposal we outlined a subsampling plan where we get this data for 100 blocks/ecoregion stratified by fire occurrence (1 block ~ 50,000 ha). I am open to a different approach but maybe we don't need lightning across the entire spatial domain?
-- 30m = yes
-- coordinate system = Canada Albers Equal Area Conic projection (ESRI:102001)

Notes on Lightning Data
-- Data have been clipped to show only those strikes that fall within 250 km of Alaska, Yukon Territory, British Columbia, and Northwest Territories
# Packages

```{r}
library(tidyverse)
library(sf)
library(lubridate)
```

# Data
```{r}
shp = st_read("../data/lightningData/CurrentYearLightning_SHP/TOA_STRIKES.shp", "TOA_STRIKES")
```

```{r}
txt = read_csv("../data/lightningData/CurrentYearLightning_TXT/TOA_STRIKES.txt")
```
```{r}
historical = read_csv("../data/lightningData/Historical_Lightning_as_txt/Historical_Lightning_2012_2023_TOA_AlaskaAlbersNAD83.txt")
```
```{r}
usa = st_read("../data/lightningData/cb_2018_us_state_20m/cb_2018_us_state_20m.shp", "cb_2018_us_state_20m")
```

## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/spatialData/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```

# Dataframes
```{r}
shp
```

```{r}
txt
```

```{r}
historical
```

```{r}
usa
```

```{r}
na = wrld %>% filter(FORMAL_EN %in% c("Canada", "United States of America"))
```
# Explore types of data
```{r}
unique(historical$STROKETYPE)
unique(historical$POLARITY)
```

# Get AK boundary
```{r}
ak = usa %>% filter(STUSPS=="AK")
```

# Split Date time column into two
```{r}
( historical1 = historical %>% 
  separate(LOCALDATETIME, into = c("LocalDate", "LocalTime"), sep = " (?=[^ ]+$)", remove = FALSE) %>%
  mutate(LocalDate = mdy(LocalDate)) %>%
  dplyr::mutate(year = lubridate::year(LocalDate), 
                month = lubridate::month(LocalDate), 
                day = lubridate::day(LocalDate))
)
```



# Quick view Plots
```{r}
plot(st_geometry(shp))
```

```{r}
plot(st_geometry(ak))
```


# Create points shapefile for Historic Lightning

## Function to make points
```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("LONGITUDE","LATITUDE"), crs = 4326, remove = FALSE)
}
```

## Drop Lat Lon NA and convert to shapefile 
```{r}
( historical2 = historical1 %>% df_to_sf() )
```

## write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(historical2, "../outputs/spatialDataCreate/HistoricalLightning.shp", driver="ESRI Shapefile")
```

# Change projections
roads <- st_transform(roads, crs = 4326)
EPSG:3338
```{r}
ls24 = st_transform(shp, crs=3338)
ak1 = st_transform(ak, crs=3338)
na1 = st_transform(na, crs=3338)
historical3 = st_transform(historical2, crs=3338)
```

# FIlter Historical by summer month and years
```{r}
historical13 = historical3 %>% filter(year==2013) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical14 = historical3 %>% filter(year==2014) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical15 = historical3 %>% filter(year==2015) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical16 = historical3 %>% filter(year==2016) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical17 = historical3 %>% filter(year==2017) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical18 = historical3 %>% filter(year==2018) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical19 = historical3 %>% filter(year==2019) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical20 = historical3 %>% filter(year==2020) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical21 = historical3 %>% filter(year==2021) %>% filter(month %in% c(5, 6, 7, 8, 9))
historical22 = historical3 %>% filter(year==2022) %>% filter(month %in% c(5, 6, 7, 8, 9))
```
# Maps

## Map area zoom
```{r}
#### zoom to kamchatka, different projection ####

zoom_to <- c(-150.46, 64.93)  # ~ center of Kamchatka
zoom_level <- 3.5

# Lambert azimuthal equal-area projection around center of interest
#target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.5)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = 3338)
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = 3338)

```

## Map Current 2024
Current Lightning is available as a shapefile for download. 

   geom_sf(data = wwf, aes(fill = ECO_NAME, color=ECO_NAME, size=prct),  pch=21, stroke =1, alpha=.75) +
  geom_sf_text(data = wwf, aes(label = prct), size=2) +
    scale_fill_manual(values=ecoPalette) +
  scale_color_manual(values=c("black", "black", "black","black","black","black","black","black","black","black","black","black","black")) +
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = ls24, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke Type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2013
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical13, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2014
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical14, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2015
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical15, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2016
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical16, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2017
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical17, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2018
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical18, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2019
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical19, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2020
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical20, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2021
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical21, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```

## Map 2022
May-September
```{r}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
    geom_sf(data = historical22, aes(fill=STROKETYPE, color=STROKETYPE), pch=21, stroke =1, alpha=.25) +
  scale_fill_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  scale_color_manual(values=c('#fc8d59', '#998ec3'), labels  = c("Cloud stroke", "Ground stroke"),
                      name = "Stroke type") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```












