---
title: "GLD Lightning data"
author: "Anna Talucci"
date: "2024-03-07"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview

GLD Data 2016-2022

**What this script does**
This script 


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

# Data
# Data

```{r}
eco = st_read("../data/spatialData/dire_ecoregs_caea/dire_ecoregs_caea.shp", "dire_ecoregs_caea")
```


## vector of data files
original data in "../data/GoogleDriveCsvFolder/"
```{r}
f <- list.files(path = "../data/lightningData/GLD/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Get Eco Projection

```{r}
st_crs(eco)
( ecozoneProjection = st_crs(eco)$proj4string )
```



# Functions

## Function to read in dataframe


## Function to clean
# Function to clean

df %>%
   separate(elements, into = c('element1', 'element2'),
          sep = '\\s*and\\s*')

```{r}
clean <- function(df){ 
  df %>% 
    mutate(eventDate = substr(time,1,10)) %>% 
    mutate(eventTime = substr(time,12,23)) %>%
    mutate(eventDate = ymd(eventDate)) %>%
    mutate(year = lubridate::year(eventDate), 
                month = lubridate::month(eventDate), 
                day = lubridate::day(eventDate)) 
}

```

## Function to make points

```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)
}
```





# View data frames

## 2016
f1 = read.csv(f[1], header = TRUE) #january
f2 = read.csv(f[2], header = TRUE) #february
f3 = read.csv(f[3], header = TRUE) #march
f3 = read.csv(f[3], header = TRUE) #march
```{r}
f5 = read.csv(f[5], header = TRUE) # may
f6 = read.csv(f[6], header = TRUE) # june
f7 = read.csv(f[7], header = TRUE) # july
f8 = read.csv(f[8], header = TRUE) # August
f9 = read.csv(f[9], header = TRUE) # september
```

f10 = read.csv(f[10], header = TRUE) #october
f11 = read.csv(f[11], header = TRUE) #november
f12 = read.csv(f[12], header = TRUE) #december

```{r}
mayToSept2016 = bind_rows(f5, f6, f7, f8, f9)
```

```{r}
( clean2016 = clean(mayToSept2016) )
```

## Clean to point
```{r}
( gld2016 = clean2016 %>% df_to_sf() )
```
## Reproject To Project Projection

```{r}
gld2016EcoProj = st_transform(gld2016, crs=ecozoneProjection)
```

## Write to shapefile

```{r}
f1 %>% mutate(eventDate = substr(time,1,10))%>% mutate(eventTime = substr(time,12,23))
```
```{r}
( clean_f1 = clean(f1) )
```


## 2017

```{r}
read.csv(f[13], header = TRUE) 
read.csv(f[14], header = TRUE) 
read.csv(f[15], header = TRUE) 
read.csv(f[16], header = TRUE) 
read.csv(f[17], header = TRUE) 
read.csv(f[18], header = TRUE) 
```

```{r}
read.csv(f[19], header = TRUE) 
read.csv(f[20], header = TRUE) 
read.csv(f[21], header = TRUE) 
read.csv(f[22], header = TRUE) 
read.csv(f[23], header = TRUE) 
read.csv(f[24], header = TRUE) 
```

## 2018

```{r}
read.csv(f[25], header = TRUE) 
read.csv(f[26], header = TRUE) 
read.csv(f[27], header = TRUE) 
read.csv(f[28], header = TRUE) 
read.csv(f[29], header = TRUE) 
read.csv(f[30], header = TRUE) 
```

```{r}
read.csv(f[31], header = TRUE) 
read.csv(f[32], header = TRUE) 
read.csv(f[33], header = TRUE) 
read.csv(f[34], header = TRUE) 
read.csv(f[35], header = TRUE) 
read.csv(f[36], header = TRUE) 
```

## 2019

```{r}
read.csv(f[37], header = TRUE) 
read.csv(f[38], header = TRUE) # No data??
read.csv(f[39], header = TRUE) 
read.csv(f[40], header = TRUE) 
read.csv(f[41], header = TRUE) 
read.csv(f[42], header = TRUE) 
```

```{r}
read.csv(f[43], header = TRUE) 
read.csv(f[44], header = TRUE) 
read.csv(f[45], header = TRUE) 
read.csv(f[46], header = TRUE) 
read.csv(f[47], header = TRUE) 
read.csv(f[48], header = TRUE) 
```

## 2020

```{r}
read.csv(f[49], header = TRUE) 
read.csv(f[50], header = TRUE) 
read.csv(f[51], header = TRUE) # march 2 data points
read.csv(f[52], header = TRUE) # april no data?
read.csv(f[53], header = TRUE) 
read.csv(f[54], header = TRUE) 
```

```{r}
read.csv(f[55], header = TRUE) 
read.csv(f[56], header = TRUE) 
read.csv(f[57], header = TRUE) 
read.csv(f[58], header = TRUE) 
read.csv(f[59], header = TRUE) 
read.csv(f[60], header = TRUE) 
```

## 2021

```{r}
read.csv(f[61], header = TRUE) 
read.csv(f[62], header = TRUE) 
read.csv(f[63], header = TRUE) 
read.csv(f[64], header = TRUE) 
read.csv(f[65], header = TRUE) 
read.csv(f[66], header = TRUE) 
```

```{r}
read.csv(f[67], header = TRUE) 
read.csv(f[68], header = TRUE) 
read.csv(f[69], header = TRUE) 
read.csv(f[70], header = TRUE) 
read.csv(f[71], header = TRUE) 
read.csv(f[72], header = TRUE) 
```

## 2022

```{r}
read.csv(f[73], header = TRUE) 
read.csv(f[74], header = TRUE) 
read.csv(f[75], header = TRUE) 
read.csv(f[76], header = TRUE) 
read.csv(f[77], header = TRUE) 
read.csv(f[78], header = TRUE) 
```

```{r}
read.csv(f[79], header = TRUE) 
read.csv(f[80], header = TRUE) 
read.csv(f[81], header = TRUE) 
read.csv(f[82], header = TRUE) 
read.csv(f[83], header = TRUE) 
read.csv(f[84], header = TRUE) 
```
```{r}
names(read.csv(f[1], header = TRUE) )
names(read.csv(f[84], header = TRUE) )
```
# Combine into a single dataframe
```{r}
listToDf <- purrr::map(f,
function(file){
df <- read.csv(file)

current_names <- names(df)

# find what names are missing
# do mutate(missing_names = NA)

return(df)
}
)
```

```{r}
gldDf = bind_rows(listToDf) 
```

```{r}
head(gldDf)
```


# Split Date time column into two
