---
title: "GLD Lightning data"
author: "Anna Talucci"
date: "2024-03-07"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview

GLD Data 2016-2022

**What this script does**
This script 


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

# Data
# Data

```{r}
eco = st_read("../data/spatialData/dire_ecoregs_caea/dire_ecoregs_caea.shp", "dire_ecoregs_caea")
```


## vector of data files
original data in "../data/GoogleDriveCsvFolder/"
```{r}
f <- list.files(path = "../data/lightningData/GLD/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

```{r}
( list2016 = f[c(1:12)] )
( list2017 = f[c(13:24)] )
( list2018 = f[c(25:36)] )
( list2019 = f[c(37:48)] )
( list2020 = f[c(49:60)] )
( list2021 = f[c(61:72)] )
( list2022 = f[c(73:84)] )

```

## 2012-2015

```{r}
txt2012 = read_csv("../data/lightningData/GLD360/GLD360_2012.txt")
txt2013 = read_csv("../data/lightningData/GLD360/GLD360_2013.txt")
txt2014 = read_csv("../data/lightningData/GLD360/GLD360_2014.txt")
txt2015 = read_csv("../data/lightningData/GLD360/GLD360_2015.txt")
```
```{r}
txt2012
txt2013
txt2014
txt2015
```

## Get Eco Projection

```{r}
st_crs(eco)
( ecozoneProjection = st_crs(eco)$proj4string )
```



# Functions

## Read into DF and Clean

```{r}
clean <- function(df){ 
  df %>% 
    map_dfr(read_csv) %>% 
  separate(time, c("date", "time"), sep = " ") %>%  
  mutate(date = ymd(date)) %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date)) %>%
    filter(month %in% c(4,5,6,7, 8, 9, 10)) %>%
    filter(cloud=="FALSE") %>%
    select(date:latitude, year:day)
  
}
```

```{r}
clean360 <- function(df){ 
  df %>%  
  mutate(Date = ymd(Date)) %>%
  dplyr::mutate(year = lubridate::year(Date), 
                month = lubridate::month(Date), 
                day = lubridate::day(Date)) %>%
    filter(month %in% c(4,5,6,7, 8, 9, 10)) %>%
    select(-Peak_current_kA) %>%
    rename(date=Date, time=Time, latitude=Latitude, longitude=Longitude)
  
}
```

# Clean

## 2016-2022

```{r}
( df2016 = list2016 %>% clean() )
( df2017 = list2017 %>% clean() )
( df2018 = list2018 %>% clean() )
( df2021 = list2021 %>% clean() )
( df2022 = list2022 %>% clean() )
```

```{r}
read_csv(list2019[12])
```
```{r}
list2019
```

```{r}
( df2019 = list2019[4:10] %>% clean() )
```
```{r}
max(df2019$month)
```
```{r}
read_csv(list2020[4])
```

```{r}
( df2020 = list2020[5:10] %>% clean() )
```

## 2012-2015

```{r}
( df2012 = txt2012 %>% clean360() )
( df2013 = txt2013 %>% clean360() )
( df2014 = txt2014 %>% clean360() )
( df2015 = txt2015 %>% clean360() )
```

## Combine

### 2016-2022

```{r eval=FALSE, include=FALSE}
combine = bind_rows(df2016, df2017, df2018, df2019, df2020, df2021, df2022)
```

```{r eval=FALSE, include=FALSE}
combine
```

### 2012-2015

```{r eval=FALSE, include=FALSE}
combine360 = bind_rows(df2012, df2013, df2014, df2015)
```

```{r eval=FALSE, include=FALSE}
combine360
```

## Function to make points

```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)
}
```

## to point
```{r}
df2012pts = df2012 %>% df_to_sf() 
df2013pts = df2013 %>% df_to_sf() 
df2014pts = df2014 %>% df_to_sf() 
df2015pts = df2015 %>% df_to_sf() 
df2016pts = df2016 %>% df_to_sf() 
df2017pts = df2017 %>% df_to_sf() 
df2018pts = df2018 %>% df_to_sf() 
df2019pts = df2019 %>% df_to_sf() 
df2020pts = df2020 %>% df_to_sf() 
df2021pts = df2021 %>% df_to_sf() 
df2022pts = df2022 %>% df_to_sf() 
```

```{r eval=FALSE, include=FALSE}
( gldpts = combine %>% df_to_sf() )
```

```{r eval=FALSE, include=FALSE}
( gld360pts = combine360 %>% df_to_sf() )
```

## Reproject To Project Projection
```{r}
df2012ptsProj = st_transform(df2012pts, crs=ecozoneProjection)
df2013ptsProj = st_transform(df2013pts, crs=ecozoneProjection)
df2014ptsProj = st_transform(df2014pts, crs=ecozoneProjection)
df2015ptsProj = st_transform(df2015pts, crs=ecozoneProjection)
df2016ptsProj = st_transform(df2016pts, crs=ecozoneProjection)
df2017ptsProj = st_transform(df2017pts, crs=ecozoneProjection)
df2018ptsProj = st_transform(df2018pts, crs=ecozoneProjection)
df2019ptsProj = st_transform(df2019pts, crs=ecozoneProjection)
df2020ptsProj = st_transform(df2020pts, crs=ecozoneProjection)
df2021ptsProj = st_transform(df2021pts, crs=ecozoneProjection)
df2022ptsProj = st_transform(df2022pts, crs=ecozoneProjection)
```

```{r eval=FALSE, include=FALSE}
gldEcoProj = st_transform(gldpts, crs=ecozoneProjection)
```

```{r eval=FALSE, include=FALSE}
gld360EcoProj = st_transform(gld360pts, crs=ecozoneProjection)
```

## write to shapefile

### shapefile

```{r eval=FALSE, include=FALSE}
st_write(gldEcoProj, "../outputs/spatialDataCreate/shapefile/GLD_2016_2022_20241127.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(gld360EcoProj, "../outputs/spatialDataCreate/shapefile/GLD360_2012_2015_20241127.shp", driver="ESRI Shapefile")
```

### geopackage
```{r eval=FALSE, include=FALSE}
st_write(gldEcoProj, "../outputs/spatialDataCreate/geopackage/GLD_2016_2022_20241127.gpkg")
```

```{r eval=FALSE, include=FALSE}
st_write(gld360EcoProj, "../outputs/spatialDataCreate/geopackage/GLD360_2012_2015_20241127.gpkg")
```

### geopackage individual years

```{r}
st_write(df2012ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2012.gpkg")
```

```{r}
st_write(df2013ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2013.gpkg")
```

```{r}
st_write(df2014ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2014.gpkg")
```

```{r}
st_write(df2015ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2015.gpkg")
```

```{r}
st_write(df2016ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2016.gpkg")
```

```{r}
st_write(df2017ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2017.gpkg")
```

```{r}
st_write(df2018ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2018.gpkg")
```

```{r}
st_write(df2019ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2019.gpkg")
```

```{r}
st_write(df2020ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2020.gpkg")
```

```{r}
st_write(df2021ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2021.gpkg")
```

```{r}
st_write(df2022ptsProj, "../outputs/spatialDataCreate/geopackage/gldByYear/GLD2022.gpkg")
```