---
title: "AK Lightning Data"
author: "Anna Talucci"
date: "2024-07-03"
output: html_document
---
```{r clearEnvironment, include=FALSE}
rm(list=ls())
```

# Overview

Explore Alaska Lightning data. Data for download within the zipped files on the [site](https://fire.ak.blm.gov/predsvcs/maps.php).

Lightning data to be used by Xanthe to evaluate wildfire probability. 

Notes:
-- Temporal period: 2013 to 2022
-- Spatial domain: Ecoregions of AK boreal interior, boreal cordillera, taiga plains, and taiga shield.
-- Lightning : yes I think we want it as precise as possible.  However, in the proposal we outlined a subsampling plan where we get this data for 100 blocks/ecoregion stratified by fire occurrence (1 block ~ 50,000 ha). I am open to a different approach but maybe we don't need lightning across the entire spatial domain?
-- 30m = yes
-- coordinate system = Canada Albers Equal Area Conic projection (ESRI:102001)

Notes on Lightning Data
-- Data have been clipped to show only those strikes that fall within 250 km of Alaska, Yukon Territory, British Columbia, and Northwest Territories

**THIS SCRIPT**
Subset data for Jeremy.

# Packages

```{r}
library(tidyverse)
library(sf)
library(lubridate)
```

# Data

```{r}
eco = st_read("../outputs/spatialDataCreate/geopackage/EcozoneBuffered.gpkg")
```

```{r eval=FALSE, include=FALSE}
gld2012_2015 = st_read("../outputs/spatialDataCreate/geopackage/GLD360_2012_2015_20241127.gpkg")
```


```{r eval=FALSE, include=FALSE}
gld2016_2022 = st_read("../outputs/spatialDataCreate/geopackage/GLD_2016_2022_20241127.gpkg")
```
## Individual years
```{r}
gld2012 = st_read("../outputs/spatialDataCreate/geopackage/gldByYear/GLD2012.gpkg")

gld2013 = st_read("../outputs/spatialDataCreate/geopackage/gldByYear/GLD2013.gpkg")

gld2014 = st_read("../outputs/spatialDataCreate/geopackage/gldByYear/GLD2014.gpkg")
```

```{r}
gld2015 = st_read("../outputs/spatialDataCreate/geopackage/gldByYear/GLD2015.gpkg")
```


## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/spatialData/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```

# Dataframes

```{r eval=FALSE, include=FALSE}
gld2012_2015
```

```{r eval=FALSE, include=FALSE}
gld2016_2022
```

# Intersect

```{r eval=FALSE, include=FALSE}
join2012_2015 = st_join(gld2012_2015, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
```

```{r eval=FALSE, include=FALSE}
join2016_2022 = st_join(gld2016_2022, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
```

## Individual Years

```{r}
join2012 = st_join(gld2012, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
join2013 = st_join(gld2013, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
join2014 = st_join(gld2014, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
join2015 = st_join(gld2015, eco, largest=TRUE) %>% drop_na(NA_L2NAME)
```

# Check Data

```{r eval=FALSE, include=FALSE}
join2012_2015
join2016_2022
```

```{r eval=FALSE, include=FALSE}
min(join2012_2015$year)
max(join2012_2015$year)
min(join2016_2022$year)
max(join2016_2022$year)
```
# Remove AK Data

```{r eval=FALSE, include=FALSE}
head(join2012_2015)
unique(join2012_2015$NA_L2NAME)
```

# Join and Select for Canada

## Subset for North America 

```{r}
na = wrld %>% filter(FORMAL_EN %in% c("Canada", "United States of America"))
```



## Define Projection

```{r}
st_crs(eco)
( ecozoneProjection = st_crs(eco)$proj4string )
```

## Reproject

```{r}
na1 = st_transform(na, crs=ecozoneProjection)
```

```{r}
na1
```

## Join to NA to Assign Country and select for Canada

```{r eval=FALSE, include=FALSE}
canada_2012_2015 = st_join(join2012_2015, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")
```

```{r eval=FALSE, include=FALSE}
canada_2016_2022 = st_join(join2016_2022, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")
```

## individual years

```{r}
canada_2012 = st_join(join2012, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")

canada_2013 = st_join(join2013, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")

canada_2014 = st_join(join2014, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")

canada_2015 = st_join(join2015, na1, largest=TRUE) %>% filter(FORMAL_EN == "Canada")
```


## drop geometry

```{r eval=FALSE, include=FALSE}
( canada_2012_2015_csv = canada_2012_2015 %>% st_drop_geometry() )
```

```{r eval=FALSE, include=FALSE}
( canada_2016_2022_csv =canada_2016_2022 %>% st_drop_geometry() )
```

## By year
```{r}
canada_2012_csv = canada_2012 %>% st_drop_geometry() 
canada_2013_csv = canada_2013 %>% st_drop_geometry() 
canada_2014_csv = canada_2014 %>% st_drop_geometry() 
canada_2015_csv = canada_2015 %>% st_drop_geometry() 
```

# Save 

```{r}
st_write(canada_2012, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2012.gpkg", layer='lightning')
```

```{r}
st_write(canada_2013, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2013.gpkg", layer='lightning')
```

```{r}
st_write(canada_2014, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2014.gpkg", layer='lightning')
```

```{r}
st_write(canada_2015, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2015.gpkg", layer='lightning')
```

```{r}
write_csv(canada_2012_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2012.csv")
```

```{r}
write_csv(canada_2013_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2013.csv")
```

```{r}
write_csv(canada_2014_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2014.csv")
```

```{r}
write_csv(canada_2015_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2015.csv")
```



```{r eval=FALSE, include=FALSE}
st_write(canada_2012_2015, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2012_2015_20241129.gpkg", layer='lightning')
```

```{r eval=FALSE, include=FALSE}
st_write(canada_2016_2022, dsn= "../outputs/spatialDataCreate/geopackage/GLD_Above_2016_2022_20241129.gpkg", layer='lightning')
```


```{r eval=FALSE, include=FALSE}
write_csv(canada_2012_2015_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2012_2015_20241129.csv")
```

```{r eval=FALSE, include=FALSE}
write_csv(canada_2016_2022_csv,"../outputs/spatialDataCreate/csv/GLD_Above_2016_2022_20241129.csv")
```
