---
title: "AK Lightning Data"
author: "Anna Talucci"
date: "2024-07-03"
output: html_document
---


# Overview

Explore Alaska Lightning data. Data for download within the zipped files on the [site](https://fire.ak.blm.gov/predsvcs/maps.php).

Lightning data to be used by Xanthe to evaluate wildfire probability. 

Notes:
-- Temporal period: 2013 to 2022
-- Spatial domain: Ecoregions of AK boreal interior, boreal cordillera, taiga plains, and taiga shield.
-- Lightning : yes I think we want it as precise as possible.  However, in the proposal we outlined a subsampling plan where we get this data for 100 blocks/ecoregion stratified by fire occurrence (1 block ~ 50,000 ha). I am open to a different approach but maybe we don't need lightning across the entire spatial domain?
-- 30m = yes
-- coordinate system = Canada Albers Equal Area Conic projection (ESRI:102001)

Notes on Lightning Data
-- Data have been clipped to show only those strikes that fall within 250 km of Alaska, Yukon Territory, British Columbia, and Northwest Territories

**THIS SCRIPT**
Subset data for Jeremy.

# Packages

```{r}
library(tidyverse)
library(sf)
library(lubridate)
```

# Data

```{r}
eco = st_read("../data/spatialData/dire_ecoregs_caea/dire_ecoregs_caea.shp", "dire_ecoregs_caea")
```
```{r}
plot(st_geometry(eco))
```

```{r}
shp = st_read("../data/lightningData/CurrentYearLightning_SHP/TOA_STRIKES.shp", "TOA_STRIKES")
```

```{r}
txt = read_csv("../data/lightningData/CurrentYearLightning_TXT/TOA_STRIKES.txt")
```
```{r}
historical = read_csv("../data/lightningData/Historical_Lightning_as_txt/Historical_Lightning_2012_2023_TOA_AlaskaAlbersNAD83.txt")
```
```{r}
usa = st_read("../data/lightningData/cb_2018_us_state_20m/cb_2018_us_state_20m.shp", "cb_2018_us_state_20m")
```

## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/spatialData/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```

# Dataframes

```{r}
eco
```

```{r}
shp
```

```{r}
txt
```

```{r}
historical
```

```{r}
usa
```

```{r}
na = wrld %>% filter(FORMAL_EN %in% c("Canada", "United States of America"))
```

# Explore types of data
```{r}
unique(historical$STROKETYPE)
unique(historical$POLARITY)
```



# Split Date time column into two

```{r}
( historical1 = historical %>% 
  separate(LOCALDATETIME, into = c("LocalDate", "LocalTime"), sep = " (?=[^ ]+$)", remove = FALSE) %>%
  mutate(LocalDate = mdy(LocalDate)) %>%
  dplyr::mutate(year = lubridate::year(LocalDate), 
                month = lubridate::month(LocalDate), 
                day = lubridate::day(LocalDate)) %>%
  filter(month %in% c(4, 5, 6, 7, 8, 9, 10)) %>%
    filter(STROKETYPE == "GROUND_STROKE")
)
```

```{r}
unique(historical1$year)
```


# Create points shapefile for Historic Lightning

## Function to make points
```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("LONGITUDE","LATITUDE"), crs = 4326, remove = FALSE)
}
```

## Drop Lat Lon NA and convert to shapefile 
```{r}
( historical2 = historical1 %>% df_to_sf() )
```


## Change projection

EPSG:3338

### Get Eco Projection

```{r}
st_crs(eco)
( ecozoneProjection = st_crs(eco)$proj4string )
```

## Reproject

```{r}
na1 = st_transform(na, crs=ecozoneProjection)
historical3 = st_transform(historical2, crs=ecozoneProjection)
```

# Select Points in ecozone polygons

```{r}
ecoBuff = st_buffer(eco, dist = 50000, endCapStyle="ROUND")
```

```{r}
ecoBuff1 = st_buffer(ecoBuff, dist = -40000, endCapStyle="ROUND")
```

```{r}
plot(st_geometry(ecoBuff1))
```

# Intersect
```{r}
joinEcoLightning = st_join(historical3 , ecoBuff1, largest=TRUE)
```

```{r}
( joinSelect = joinEcoLightning %>% 
    drop_na(NA_L2NAME) %>%
  select(OID_:geometry)
)
  
```
```{r}
min(joinSelect$year)
max(joinSelect$year)
```
## Join and select for US Only

```{r}
( AK_2012_2022 = st_join(joinSelect, na1, largest=TRUE) %>% 
  filter(FORMAL_EN == "United States of America") %>% 
  filter(year != 2023) )
```

```{r}
min(AK_2012_2022$year)
max(AK_2012_2022$year)

```

```{r}
AK_2012_2022 %>% filter(year==2013)

```
2013 has 34,822

## write to shapefile

```{r eval=FALSE, include=FALSE}
st_write(joinSelect, "../outputs/spatialDataCreate/AkLightning12_23.shp", driver="ESRI Shapefile")
```

```{r}
st_write(AK_2012_2022, dsn= "../outputs/spatialDataCreate/AkLightning2012_2022_20240915.gpkg", layer='lightning')
```

## write csv
```{r}
( joinSelectDf = joinSelect %>% st_drop_geometry() )
```
```{r}
write_csv(joinSelectDf ,"../outputs/spatialDataCreate/AkLightning12_23.csv")
```

```{r}
( AK_2012_2022_csv = AK_2012_2022 %>% st_drop_geometry() )
```
```{r}
write_csv(AK_2012_2022_csv ,"../outputs/spatialDataCreate/csv/AkLightning2012_2023_20240915.csv")
```

# Read geopackage

```{r}
# read in layers to environment, suppress messages with quiet=TRUE
akgpkg <- st_read(dsn =  "../outputs/spatialDataCreate/AkLightning13_23.gpkg", layer='lightning')
```
```{r}
akgpkg
```

# Maps

## Map area zoom
```{r zoomMap}
#### zoom to kamchatka, different projection ####

zoom_to <- c(-130.46, 64.93)  # ~ center of Kamchatka
zoom_level <- 3

# Lambert azimuthal equal-area projection around center of interest
#target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.5)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = ecozoneProjection)
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = ecozoneProjection)

```

```{r map, echo=FALSE}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
  
    geom_sf(data = joinSelect, aes(fill='#998ec3', color='#998ec3'), pch=21, stroke =1, alpha=.25) +
 geom_sf(data = eco, fill = NA, colour="black") +
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", linewidth = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```